### 2018-06-15 Paper/method.tex
changes on bg from yesterday ...
..

In this section, we outline the underlying framework for \emph{adaptively}
constructing a reduced-space surrogate (RSS) using sensitivity analysis.  The
proposed methodology is described as adaptive since the RSS is constructed only
in situations where it is expected to yield computational dividend as discussed
further below.  The term reduced-space implies that the surrogate is
constructed in a subspace that sufficiently captures the uncertainty in the
model output. We begin by outlining an algorithm for parameter screening
to assess the importance of individual parameters for potential dimension
reduction and construction of an RSS. The overall adaptive framework 
which incorporates parameter screening as an integral step is thereafter
presented. Finally, we present metrics used for assessing the
convergence and accuracy of the RSS followed by a brief discussion on
salient features of the proposed framework.

\textbf{Parameter screening.}
Dimension reduction is guided by global sensitivity analysis (GSA).  A typical
approach to GSA involves estimating the Sobol sensitivity
indices~\cite{Sobol:2001} in the case of non-linear mathematical models.
However, as discussed earlier in sections~\ref{sec:intro} and~\ref{sec:bg},
determining accurate estimates of Sobol sensitivity indices typically requires
tens of thousands of model evaluations. Consequently, the exercise becomes
prohibitive for expensive-to-evaluate models.  In the proposed framework, we
adopt a novel approach for constructing an RSS based on estimating the
upper-bound $\widehat{\mathcal{C}_i\mu_i}$, given in~\eqref{eq:bound}, on 
Sobol total effect index ($\mathcal{T}(\theta_i)$) for each 
parameter, $\theta_i$, and use it as a metric to screen 
parameters that are relatively unimportant.

The metric is initially computed using model evaluations at say, $n_1$ samples
in the full space. 
In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically,
we could use finite-difference to approximate its value as shown in
Eq.~\ref{eq:partial}. However, as discussed earlier in~\ref{sub:dgsm},
estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in case
the derivative is available analytically. Hence, computational effort when
using finite differences is expected to increase by a factor of $(N_p+1)$.
Another possible approach, suitable in certain cases, is the use of
adjoint-based gradient computation. In particular, for PDE-based
problems with scalar QoIs, but high-dimensional uncertain parameters, 
adjoint-based gradient computation is much more efficient than finite differences.
In the adjoint approach, each gradient evaluation requires 
a solution of the state equation (forward solve) and that of the 
corresponding adjoint equation; 
see e.g.,~\cite{jameson1988aerodynamic,gunzburger2003perspectives,Borzi2011}.
The adjoint method, however, requires the availability of an adjoint solver.

Based on the associated metric value, an initial
rank ($\mathcal{R}_i^{old}$) is assigned to each parameter. At each iteration,
a new set of Latin Hypercube samples is generated and corresponding model
output at each sample point is computed. The new set of model evaluations
combined with prior evaluations is used to update parameter ranks. Additionally,
deviation in the derivative-based sensitivity measure between successive iterations
normalized by the measure in the previous iteration is recorded for each parameter.
The iterative process is continued until parameter ranks between successive iterations
are observed to be consistent as well as the maximum deviation among all parameters
($\Delta\mu_s$) is below a certain tolerance ($\tau$). The amount of computational
effort associated with the screening process is limited
by the choice of maximum number of iterations, $s_\text{max}$. 
Key inputs to the screening procedure
are as follows: (1) 
a limiting value $\tau$ of the maximum relative change in the sensitivity
measure between successive iterations; (2) 
a limiting ratio $\tau_\text{screen}$ of the sensitivity metric 
relative to its maximum value;
(3) a real number $\beta \in (0, 1)$ to 
guide the number of new samples $\lceil \beta n_1 \rceil$ 
at each iteration ($\lceil \beta n_1
\rceil$ 
is the smallest integer greater than or equal to $\beta n_1$); 
(4) a set of samples $\{ \bm{\theta}_k \}_{k = 1}^{n_0}$ 
for the initial screening step in
the algorithm and the corresponding gradient evaluations 
$\{ \bm{g}^k \}_{k=1}^{n_0}$, where
$\bm{g}^k = \nabla_{\bm{\theta}} G(\bm{\theta}_k)$. 
The outputs are the set of active
indices $\mathcal{I}_\text{active}$ corresponding to the 
\emph{important parameters}, the total number of available model 
evaluations $N_\text{total}$, and the enriched set of gradient evaluations 
$\{ \bm{g}^k\}_{k=1}^{N_\text{total}}$. 
A general methodology for parameter screening is provided below in
Algorithm~\ref{alg:screen}.
%\clearpage

@@

We begin by setting aside computational resources for constructing a cross-validation
test suite to be used for assessing the accuracy of the resulting surrogate.
Naturally, the resources allocated for this purpose depend upon the application
as well as total amount of available resources. 
The set of required inputs for parameter screening are initialized,
and model evaluations at $n_1$ random samples in the full-space are computed.
These evaluations are used to construct a surrogate in the full-space (FSS)
using regression-based techniques. If the surrogate is found to be sufficiently
accurate for the given application, the process is terminated. However, it is
likely that a full-space surrogate constructed using a 
small number of model evaluations would not provide a 
faithful representation of the input-output relationship.

@@

\textbf{Assessment of the surrogate.}
To assess convergence of the resulting surrogate, one could estimate the
leave-one-out cross validation error as follows:
\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee
where $N_l$ is the number of training points,
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$ is
the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash
i}}$ is the PC surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$
model realization.  From~\eqref{eq:loo}, it appears that $N_l$ PCEs are needed
to evaluate $\epsilon_{\mbox{\tiny LOO}}$.  However, in practice a modified
formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009}, independent
of $G^{\mbox{\tiny {PC}\textbackslash i}}$ is used; for an easy reference,
see~\cite[Eq.~(1.27)]{Marelli:2014}.  Accuracy of the surrogate could also 
be assessed by evaluating the relative $\ell_2$-norm of the difference in
predictions between the model and the surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:
\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee
where $N_v$ is the number of sampling points in the full parameter space at
which model evaluations are available. In the case of an RSS, it is the
augmented set of model evaluations used for validation and screening.  
%
Accuracy of the surrogate could be further investigated by comparing
probability density functions (PDFs) of the model output based on model
evaluations in the full parameter space and the RSS predictions corresponding
to a large number of samples (say, 10$^6$ for a high-dimensional input space).
However, in realistic problems
involving complex, compute-intensive simulations, constructing the PDF based on model
evaluations would be infeasible.  A practical alternative would be to compare
a histogram based on sparse model evaluations with the surrogate-based PDF in
order to gain some insight into the quality of the comparison.

\textbf{Remarks.}
The amount of computational effort associated with the presented methodology
can be mainly attributed to two steps: I.~Parameter Screening, and 
II.~Constructing a converged RSS. Computational gains are realized in situations
where constructing the surrogate in the full parameter space is more expensive
than the combined cost associated with these steps. Determining the optimal
allocation of computational resources for these steps, however, is not possible
a priori. Hence, in the proposed framework, we exploit the set of model
evaluations used in parameter screening to simultaneously construct the FSS
while keeping a track of its accuracy using the validation test suite. This
would help address situations where significant dimension reduction is not
possible, and hence, constructing the RSS might result in a computational
disadvantage. We suggest using a small number of samples in the initial
screening step (say, $n_1$ = 5) and a relatively large $\tau$ (say,
$\mathcal{O}(10^{-1})$) as a starting point with possible reduction in $\tau$
during subsequent screenings.

The applicability of the proposed framework depends upon the choice of the
model output.  Since the screening metric involves computation of partial
derivatives in the full parameter space, the output must exhibit differentiable
dependence on each parameter. It is therefore likely that for a given
application involving multiple outputs, the RSS can only be constructed for a
selected few, using the approach presented above.  Hence, it is important to
assess the nature of the input-output relationship for a given model prior to
implementing the present framework.

Using the framework proposed in this section, we aim to construct a reliable
surrogate in the most efficient manner within the constraints of the computational
budget. However, it is likely that for a given application, the RSS is not
found to be sufficiently accurate. In such a scenario, we suggest enriching the
set of important inputs by incorporating the least unimportant model input
as determined after a series of screening steps, and re-constructing the RSS. 
This process could be repeated depending upon the availability of resources
and the desired accuracy of the surrogate.

>>>>

In this section, we outline the underlying framework for \emph{adaptively}
constructing a reduced-space surrogate (RSS) using sensitivity analysis.  The
proposed methodology is described as adaptive since the RSS is constructed only
in situations where it is expected to yield computational dividend as discussed
further below.  The term reduced-space implies that the surrogate is
constructed in a reduced parameter space that sufficiently captures the uncertainty in the
model output. We begin by outlining an algorithm for parameter screening
to assess the importance of individual parameters for potential dimension
reduction and construction of an RSS. The overall adaptive framework 
that incorporates parameter screening as an integral step is thereafter
presented. Finally, we present metrics used for assessing the
convergence and accuracy of the RSS followed by a brief discussion on
salient features of the proposed framework.

\textbf{Parameter screening.}
In the proposed framework, we
adopt a novel approach for constructing an RSS based on estimating the
upper-bound $\widehat{\mathcal{C}_i\mu_i}$, given in~\eqref{eq:bound}, on 
Sobol total effect index ($\mathcal{T}(\theta_i)$) for each 
parameter, $\theta_i$, and use it as a metric to screen 
parameters that are relatively unimportant.

%The metric is initially computed using model evaluations at say, $n_1$ samples
%in the full space. 
%In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically,
%we could use finite-difference to approximate its value as shown in
%Eq.~\ref{eq:partial}. However, as discussed earlier in~\ref{sub:dgsm},
%estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires
%model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in case
%the derivative is available analytically. Hence, computational effort when
%using finite differences is expected to increase by a factor of $(N_p+1)$.
%Another possible approach, suitable in certain cases, is the use of
%adjoint-based gradient computation. In particular, for PDE-based
%problems with scalar QoIs, but high-dimensional uncertain parameters, 
%adjoint-based gradient computation is much more efficient than finite differences~\cite{Griewank:2008}.
%In the adjoint approach, each gradient evaluation requires 
%a solution of the state equation (forward solve) and that of the 
%corresponding adjoint equation; 
%see e.g.,~\cite{jameson1988aerodynamic,gunzburger2003perspectives,Borzi2011}.
%The adjoint method, however, requires the availability of an adjoint solver.

An initial set of say, $n_1$ samples is used to obtain a rough estimate
of the metric.  
Based on the associated metric value, an initial
rank ($\mathcal{R}_i^{old}$) is assigned to each parameter. At each iteration,
a new set of samples is generated based on the joint probability distribution
of $\bm{\theta}$ and corresponding model
output at each sample point is computed. The new set of model evaluations
combined with prior evaluations is used to update parameter ranks. Additionally,
deviation in the derivative-based sensitivity measure between successive iterations
normalized by the measure in the previous iteration is recorded for each parameter.
The iterative process is continued until parameter ranks between successive iterations
are observed to be consistent as well as the maximum deviation among all parameters
($\Delta\mu_s$) is below a certain tolerance ($\tau$). The amount of computational
effort associated with the screening process is limited
by the choice of maximum number of iterations, $s_\text{max}$.

Key inputs to the screening procedure
are as follows: (1) 
a limiting value $\tau$ of the maximum relative change in the sensitivity
measure between successive iterations; (2) 
a limiting ratio $\tau_\text{screen}$ of the sensitivity metric 
relative to its maximum value;
(3) a real number $\beta \in (0, 1)$ to 
guide the number of new samples $\lceil \beta n_1 \rceil$ 
at each iteration ($\lceil \beta n_1
\rceil$ 
is the smallest integer greater than or equal to $\beta n_1$); 
(4) a set of samples $\{ \bm{\theta}_k \}_{k = 1}^{n_0}$ 
for the initial screening step in
the algorithm and the corresponding gradient evaluations 
$\{ \bm{g}^k \}_{k=1}^{n_0}$, where
$\bm{g}^k = \nabla_{\bm{\theta}} G(\bm{\theta}_k)$. 
The outputs are the set of active
indices $\mathcal{I}_\text{active}$ corresponding to the 
\emph{important parameters}, the total number of available model 
evaluations $N_\text{total}$, and the enriched set of gradient evaluations 
$\{ \bm{g}^k\}_{k=1}^{N_\text{total}}$. 
A general methodology for parameter screening is provided below in
Algorithm~\ref{alg:screen}.
%\clearpage

@@

We begin by allocating computational resources for constructing a cross-validation
test suite to be used for assessing the accuracy of the resulting surrogate.
Naturally, the resources allocated for this purpose depend upon the application
as well as total amount of available resources. 
The set of required inputs for parameter screening are initialized,
and model evaluations at $n_1$ random samples in the full-space are computed.
These evaluations are used to construct a surrogate in the full-space (FSS)
using regression-based techniques. If the surrogate is found to be sufficiently
accurate for the given application, the process is terminated. However, it is
likely that a full-space surrogate constructed using a 
small number of model evaluations would not provide a 
faithful representation of the input-output relationship.

@@

\textbf{Assessment of the surrogate.}
To assess convergence of the resulting surrogate, one could estimate the
leave-one-out cross validation error as follows:
\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2},
\label{eq:loo}
\ee
where $N_l$ is the number of training points,
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$ is
the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash
i}}$ is the PC surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$
model realization.  From~\eqref{eq:loo}, it appears that $N_l$ PCEs are needed
to evaluate $\epsilon_{\mbox{\tiny LOO}}$.  However, in practice a modified
formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009}, independent
of $G^{\mbox{\tiny {PC}\textbackslash i}}$ is used; for an easy reference,
see~\cite[Eq.~(1.27)]{Marelli:2014}.  Accuracy of the surrogate could also 
be assessed by evaluating the relative $\ell_2$-norm of the difference in
predictions between the model and the surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:
\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}},
\label{eq:l2}
\ee
where $N_v$ is the number of sampling points in the full parameter space at
which model evaluations are available. In the case of an RSS, it is the
augmented set of model evaluations used for validation and screening.  
%
Accuracy of the surrogate could be further investigated by comparing
probability density functions (PDFs) of the model output based on model
evaluations in the full parameter space and the RSS predictions corresponding
to a large number of samples (say, 10$^6$ for a high-dimensional input space).
However, in realistic problems
involving complex, compute-intensive simulations, constructing the PDF based on model
evaluations would be infeasible.  A practical alternative would be to compare
a histogram based on sparse model evaluations with the surrogate-based PDF in
order to gain some insight into the quality of the comparison.

\textbf{Discussion on the proposed methodology.}
The amount of computational effort associated with the presented methodology
can be mainly attributed to two steps: I.~Parameter Screening, and 
II.~Constructing a converged RSS. Computational gains are realized in situations
where constructing the surrogate in the full parameter space is more expensive
than the combined cost associated with these steps. Determining the optimal
allocation of computational resources for these steps, however, is not possible
a priori. Hence, in the proposed framework, we exploit the set of model
evaluations used in parameter screening to simultaneously construct the FSS
while keeping a track of its accuracy using the cross-validation test suite.
This would help address situations where significant dimension reduction 
is not possible, and hence, constructing the RSS might result in a 
computational
disadvantage. We suggest using a small number of samples in the initial
screening step (say, $n_1$ = 5) and a relatively large $\tau$ (say,
$\mathcal{O}(10^{-1})$) as a starting point with possible reduction in $\tau$
during subsequent screenings. Pseudo-random sampling approaches such as
Latin hypercube sampling (LHS) and quasi Monte Carlo (QMC) could be used to
generate samples in the input space.

Careful assessment and decision-making is required on whether or not to
proceed with the construction of the RSS at the end of each screening step.
The user should account for factors such as the possible degree of 
dimension reduction, accuracy of the concurrent FSS, and availability of
computational resources.

The applicability of the proposed framework depends upon the choice of the
model output.  Since the screening metric involves computation of partial
derivatives in the full parameter space, the output must exhibit differentiable
dependence on each parameter. It is therefore likely that for a given
application involving multiple outputs, the RSS can only be constructed for a
selected few, using the approach presented above.  Hence, it is important to
assess the nature of the input-output relationship for a given model prior to
implementing the present framework. Additionally, in some cases, the partial
derivative of the output with respect to each uncertain input is not available
analytically. In these cases, one could use finite difference (FD) to approximate
the gradient as illustrated in~\ref{eq:partial}. However, since FD requires
model evaluations at neighboring points, the underlying computational cost is
expected to increase by a factor, $N_p+1$; $N_p$ being the number of inputs. 
A possible, more efficient alternative to FD involves the use of adjoints for
gradient computation~\cite{Griewank:2008}. In the adjoint approach, each
gradient evaluation requires 
a solution of the state equation (forward solve) and that of the 
corresponding adjoint equation; 
see e.g.,~\cite{jameson1988aerodynamic,gunzburger2003perspectives,Borzi2011}.
The adjoint method, however, requires the availability of an adjoint solver.

Using the framework proposed in this section, we aim to construct a reliable
surrogate in the most efficient manner within the constraints of the computational
budget. However, it might be possible that for a given application, the RSS is not
found to be sufficiently accurate. In such a scenario, we suggest enriching the
set of important inputs by incorporating the least unimportant model input
as determined after a series of screening steps, and re-constructing the RSS. 
This process could be repeated depending upon the availability of resources
and the desired accuracy of the surrogate.
